---
title: "Introduction to Curve Aware"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Curve Aware}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Installing and Loading `CurveAware`
The purpose of this "package" is to provide a framework to fit age-specific infection fatality rates (IFRs) or the initial number of individuals infected ($I_0$) in an epidemic under several key assumptions. To install this package, you will need to first install the [`remotes`](https://cran.r-project.org/web/packages/remotes/index.html) package and then run the following command:   

```{r, eval=FALSE}
remotes::install_github("nickbrazeau/CurveAware")
```

As previously discussed, this package is essentially a "wrapper" of the [`drjacoby`](https://github.com/mrc-ide/drjacoby) R package, which depends on the [Rcpp](https://cran.r-project.org/web/packages/Rcpp/index.html) package. The `Rcpp` package requires that certian OS-specific dependcies are avialble, as stated on [`drjacoby`](https://github.com/mrc-ide/drjacoby) site and pasted below for ease: 

* Windows
    - Download and install the appropriate version of [Rtools](https://cran.rstudio.com/bin/windows/Rtools/) for your version of R. On installation, ensure you check the box to arrange your system PATH as recommended by Rtools
* Mac OS X
    - Download and install [XCode](http://itunes.apple.com/us/app/xcode/id497799835?mt=12)
    - Within XCode go to Preferences : Downloads and install the Command Line Tools
* Linux (Debian/Ubuntu)
    - Install the core software development utilities required for R package development as well as LaTeX by executing
    ```
    sudo apt-get install r-base-dev texlive-full
    ```

If you have any problems installing then please [raise an issue](https://github.com/nickbrazeau/CurveAware/issues) on github.

Assuming everything installed correctly, you can now load the package:

```{r, eval=FALSE}
library(CurveAware)
```


