---
title: "Introduction to Curve Aware"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Curve Aware}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, echo=FALSE}
library(ggplot2)
library(magrittr)
library(CurveAware)
```


## Introduction
The purpose of this "package" is to provide a framework to fit the infection-fatality-rate (IFR) or the initial number of individuals infected ($I_0$) in an epidemic under several key assumptions (see below). We fit our model using the Metropolis-Coupled MCMC withthe [`drjacoby`](https://mrc-ide.github.io/drjacoby/) package (_so much so that this "package" is more appropriately called a wrapper of `drjacoby`_). 

### Overview of the `CurveAware` Model 
We assume that the only data avaialble is the cumulative number of deaths at a given day $t$ during the epidemic. In addition, we assume that the epidemic is growing exponentially (_i.e._ saturation of the fraction susceptible has not been reached). We then "cast" our incidence curve on to a gamma-distributed curve that represents the delay of disease onset to death. 

```{r, results='asis', echo=FALSE, fig.align='center', fig.width=6, fig.height=3}
set.seed(48)
infxn <- data.frame(time = 1:100, It = sapply(1:100, function(x, I0, r){I0*exp(r*x)}, I0 = 50, r = 0.10))
tod <- data.frame(time = 1:15, todt = sapply(1:15, function(x, m_od, s_od){pgamma(x, shape = 1/s_od^2, scale = m_od*s_od^2)}, m_od = 10, s_od = 0.5))


#..................
# plot
#..................

infxnplot <- infxn %>% 
  ggplot() + 
  geom_line(aes(x = time, y = It), color = "#3288bd", size = 3) +
  geom_area(aes(x = ifelse(time >= 0 & time <= 100, time, 0), y = It), fill = "#3288bd", alpha = 0.5) +
  theme_minimal() +
  xlab("Time") + ylab("Num. Infected") +
  theme(
    axis.title = element_text(face = "bold", size = 16),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line(color = "#000000", size = 2)
  )
  
Xplot <- ggplot() +
  geom_text(aes(x = 1, y = 1, label = paste('bolditalic("X")')), parse = T, size = 15) + 
  xlim(0.9, 1.1) + ylim(0.9, 1.1) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank(), 
    panel.border = element_blank(), 
    strip.background = element_blank(),
  )

todplot <- tod %>% 
  ggplot() + 
  geom_line(aes(x = time, y = todt), color = "#d53e4f", size = 3) +
  geom_area(aes(x = ifelse(time >= 0 & time <= 100, time, 0), y = todt), fill = "#d53e4f", alpha = 0.5) +
  theme_minimal() +
  xlab("Time") + ylab("Prob. Onset to Death") +
  theme(
    axis.title = element_text(face = "bold", size = 16),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line(color = "#000000", size = 2)
  )
  

cowplot::plot_grid(infxnplot, Xplot, todplot, 
                   rel_heights = c(2, 0.75, 2),
                   rel_widths = c(2, 0.75, 2), 
                   nrow = 1)


```

As a result, the expected number of deaths on a given day, $t$ is a function of the total number of individuals infected on that day accounting for the offset of time infection to (potential) death.   
  
Given that there are two parts to this function (that are joint), one can either put a strong prior on the: (1) IFR and infer the $I_0$ (and the "exponential curve" of the epidemic); or (2) $I_0$ and infer the IFR. 
