---
title: "Running the Model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running the Model}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


## `COVIDCurve` Purpose
The purpose of this package is to provide a framework to fit age-specific infection fatality ratios (IFRs) during an epidemic using seroprevalence data to track cumulative infections. The model was originally designed in response to the COVID-19 pandemic (hence the name) but is broadly applicable.


```{r setup, message = F}
library(magrittr)
library(tibble)
library(ggplot2)
library(cowplot)
library(COVIDCurve)
```


## Simulating Data
Here, we will use the model's "back-end" simulator to create some data. The simulator was built "under the model" and reflects the forward process, or generative process, that underlies the model's inferential framework. 


```{r simdata}
set.seed(1234)
#............................................................
# Pieces for the Simulation
#...........................................................
# users are required to input their own "infection" curve, 
# here we will use a simple sigmoidal function with some noise
infxns <- data.frame(time = 1:200)
sig <- function(x){1 / (1 +  exp(-x))}
timevec <- seq(from = -5, to = 7, length.out = nrow(infxns))
infxns$infxns <- sig(timevec) * 5e3 + runif(n = nrow(infxns), 
                                            min = -25,
                                            max = 50)

# make up IFR values and attack rates within the three "ma" age groups 
fatalitydata <- tibble::tibble(Strata = c("ma1", "ma2", "ma3"),
                               IFR = c(0.01, 0.05, 0.1),
                               Rho = 1) # assuming uniform attack rate wrt demographic size
# population sizes
demog <- tibble::tibble(Strata = c("ma1", "ma2", "ma3"),
                        popN = c(1500000, 2250000, 1250000))

#............................................................
# Running the Simulation Function 
#...........................................................
simdat <- COVIDCurve::Agesim_infxn_2_death(
  fatalitydata = fatalitydata,
  demog = demog,
  m_od = 19.8,
  s_od = 0.85,
  curr_day = 200,
  infections = infxns$infxns,
  simulate_seroreversion = FALSE,
  sens = 0.85,
  spec = 0.95,
  sero_delay_rate = 18.3
)
```

Our simulated data is now a list with three dataframes: (1) StrataAgg_TimeSeries_Death: stratified daily deaths by age group; 
(2) Agg_TimeSeries_Death: summed daily deaths (summed across age groups); (3) StrataAgg_Seroprev: daily seroprevalences by age group.

We will now tidy up these simulated data items for the model fitting steps. 

```{r tidydata}

# here we are calculating the total proportion of deaths
# within each age group 
# proportion deaths
prop_deaths <- simdat$StrataAgg_TimeSeries_Death %>%
  dplyr::group_by(Strata) %>%
  dplyr::summarise(deaths = sum(Deaths)) %>%
  dplyr::ungroup(.) %>%
  dplyr::mutate(PropDeaths = deaths/sum(simdat$Agg_TimeSeries_Death$Deaths)) %>%
  dplyr::select(-c("deaths"))

# Serologic studies are typically conducted over a few days
# here we have assumed that two time points with 5 days on
# either side of our "midpoint".
# For each study and we will take the average over that time 

sero_days <- c(135, 160)
sero_days <- lapply(sero_days, function(x){seq(from = (x-5), to = (x+5), by = 1)})
obs_serology <- simdat$StrataAgg_Seroprev %>%
  dplyr::group_by(Strata) %>%
  dplyr::filter(ObsDay %in% unlist(sero_days)) %>%
  dplyr::mutate(serodaynum = sort(rep(1:length(sero_days), 11))) %>%
  dplyr::mutate(
    SeroPos = ObsPrev * testedN,
    SeroN = testedN ) %>%
  dplyr::group_by(Strata, serodaynum) %>%
  dplyr::summarise(SeroPos = mean(SeroPos),
                   SeroN = mean(SeroN)) %>% # seroN doesn't change
  dplyr::mutate(SeroStartSurvey = sapply(sero_days, median) - 5,
                SeroEndSurvey = sapply(sero_days, median) + 5,
                SeroPos = round(SeroPos),
                SeroPrev = SeroPos/SeroN,
                SeroLCI = NA,
                SeroUCI = NA) %>%
  dplyr::select(c("SeroStartSurvey", "SeroEndSurvey", "Strata", "SeroPos", "SeroN", "SeroPrev", "SeroLCI", "SeroUCI")) %>%
  dplyr::ungroup(.) %>%
  dplyr::arrange(SeroStartSurvey, Strata)

```

## Building Model Object
We will now make our model fitting object. 

```{r}
#......................
# put binomial data in correct format 
#......................
inputdata <- list(obs_deaths = simdat$Agg_TimeSeries_Death,
                  prop_deaths = prop_deaths,
                  obs_serology = obs_serology)

#......................
# Prior distrubitions and sampling distributions
#......................
# sens/spec
sens_spec_tbl <- tibble::tibble(name =  c("sens",  "spec"),
                                min =   c(0.5,      0.5),
                                init =  c(0.85,     0.95),
                                max =   c(1,        1),
                                dsc1 =  c(850.5,    950.5),
                                dsc2 =  c(150.5,    50.5))

# delay priors
tod_paramsdf <- tibble::tibble(name = c("mod", "sod", "sero_con_rate"),
                               min  = c(18,     0,     16),
                               init = c(19,     0.85,  18),
                               max =  c(20,     1,     21),
                               dsc1 = c(19.8,   2550,  18.3),
                               dsc2 = c(0.1,    450,   0.1))

# make param dfs
ifr_paramsdf <- tibble::tibble(name = c("ma1", "ma2",  "ma3"),
                               min  = rep(0, 3),
                               init = rep(0.2, 3),
                               max = rep(0.4, 3),
                               dsc1 = rep(0, 3),
                               dsc2 = rep(0.4, 3))

infxn_paramsdf <- tibble::tibble(name = paste0("y", 1:5),
                                 min  = rep(0, 5),
                                 init = rep(2, 5),
                                 max =  rep(10, 5),
                                 dsc1 = rep(0, 5),
                                 dsc2 = rep(10, 5))

knot_paramsdf <- tibble::tibble(name = paste0("x", 1:4),
                                min  = c(2,  50, 100, 150),
                                init = c(5,  75, 125, 175),
                                max =  c(49, 99, 149, 200),
                                dsc1 = c(2,  50, 100, 150),
                                dsc2 = c(49, 99, 149, 200))


noise_paramsdf <- tibble::tibble(name = c("ne1", "ne2", "ne3"),
                                 min  = rep(0.5, 3),
                                 init = rep(1, 3),
                                 max = rep(1.5, 3),
                                 dsc1 = rep(1, 3),
                                 dsc2 = rep(0.05, 3))

# bring together
df_params <- rbind.data.frame(ifr_paramsdf, infxn_paramsdf, noise_paramsdf, knot_paramsdf, sens_spec_tbl, tod_paramsdf)

#......................
# make model for serorev and regular
#......................
# reg
mod1 <- COVIDCurve::make_IFRmodel_age$new()
mod1$set_MeanTODparam("mod")
mod1$set_CoefVarOnsetTODparam("sod")
mod1$set_IFRparams(paste0("ma", 1:3))
mod1$set_maxMa("ma3")
mod1$set_Knotparams(paste0("x", 1:4))
mod1$set_Infxnparams(paste0("y", 1:5))
mod1$set_Noiseparams(c(paste0("ne", 1:3)))
mod1$set_Serotestparams(c("sens", "spec", "sero_con_rate"))
mod1$set_data(inputdata)
mod1$set_demog(demog)
mod1$set_paramdf(df_params)
mod1$set_rcensor_day(.Machine$integer.max) # no censoring
```

<!-- ## Running MCMC  -->

<!-- ```{r, echo=F, eval=T} -->
<!-- # "Warm-Up" MCMC -->
<!-- warmdf_params <- rbind.data.frame(list("x", 1, 1, 1)) -->
<!-- names(warmdf_params) <- c("name", "min", "max", "init") -->
<!-- warmloglike <- "SEXP loglike(Rcpp::NumericVector params, int param_i, Rcpp::List data, Rcpp::List misc) { double ret = -1.0; return Rcpp::wrap(ret);}" -->
<!-- warmlogprior <- "SEXP logprior(Rcpp::NumericVector params, int param_i, Rcpp::List misc) { double ret = -1.0; return Rcpp::wrap(ret);}" -->
<!-- warmup <- drjacoby::run_mcmc(data = list("dat" = c(1)), -->
<!--                              df_params = warmdf_params, -->
<!--                              misc = list(), -->
<!--                              loglike = warmloglike, -->
<!--                              logprior = warmlogprior, -->
<!--                              burnin = 1, -->
<!--                              samples = 1, -->
<!--                              chains = 1, -->
<!--                              rungs = 1, -->
<!--                              silent = T) -->
<!-- ``` -->

```{r run MCMC}
start <- Sys.time()
modout <- COVIDCurve::run_IFRmodel_age(IFRmodel = mod1,
                                       reparamIFR = FALSE,
                                       reparamInfxn = FALSE,
                                       reparamKnot = FALSE,
                                       burnin = 1e4,
                                       samples = 1e4,
                                       chains = 3,
                                       rungs = 1,
                                       thinning = 0,
                                       silent = TRUE)
end <- Sys.time() - start
```
Here we will run a very fast MCMC that took only `r round(end)`. However, with much larger models containing 10,000 burn-in and sampling iterations across 50 rungs, run times can exceed 24-36 hours. These large runs will require a cluster. 


## Examining the various outputs

### Posterior Summaries
```{r}
(ifr <- COVIDCurve::get_cred_intervals(IFRmodel_inf = modout, whichrung = paste0("rung", 1),
                                       what = "IFRparams", by_chain = F))

(sero <- COVIDCurve::get_cred_intervals(IFRmodel_inf = modout, whichrung = paste0("rung", 1),
                                        what = "Serotestparams", by_chain = F))
```

```{r, eval=F}
# the below posteriors are usually only needed for infetion curves
# and posterior checks
(knotspost <- COVIDCurve::get_cred_intervals(IFRmodel_inf = modout,  whichrung = paste0("rung", 1),
                                             what = "Knotparams", by_chain = F))
(infxn <- COVIDCurve::get_cred_intervals(IFRmodel_inf = modout,  whichrung = paste0("rung", 1),
                                         what = "Infxnparams", by_chain = F))

(ddelays <- COVIDCurve::get_cred_intervals(IFRmodel_inf = modout,  whichrung = paste0("rung", 1),
                                           what = "DeathDelayparams", by_chain = F))
(neparams <- COVIDCurve::get_cred_intervals(IFRmodel_inf = modout,  whichrung = paste0("rung", 1),
                                            what = "Noiseparams", by_chain = F))

```

### Visualizing Output
#### Infection Curve
Here we can downsample from the posterior to get a sense of our inferred infection curve. 
```{r, fig.width=8, fig.height=6, fig.align='center'}
curve <- COVIDCurve::draw_posterior_infxn_cubic_splines(IFRmodel_inf = modout,
                                                        whichrung = paste0("rung", 1),
                                                        by_chain = F,
                                                        by_strata = F,
                                                        dwnsmpl = 1e3)
#......................
# some extra plot items
#......................
# tidy up and make plots
liftover <- data.frame(param = c("ma1", "ma2", "ma3"),
                       Strata = c("ma1", "ma2", "ma3"))

fatalitydataplot <- fatalitydata %>%
  dplyr::left_join(liftover, ., by = "Strata")

plot1 <- ggplot() +
  geom_pointrange(data = ifr, aes(x = param, ymin = LCI, ymax = UCI, y = median, color = param)) +
  geom_hline(data = fatalitydataplot, aes(yintercept  = IFR, group = param), color = "#3182bd", linetype = "dashed", size = 1.1) +
  facet_wrap(.~param) +
  scale_color_viridis_d() +
  theme_bw()

plot2 <- curve$plotObj +
  geom_line(data = infxns, aes(x = time, y = infxns), color = "#3182bd")

cowplot::plot_grid(plot1, plot2, ncol = 1, nrow = 2)


```

#### Seroprevalence Curve
Similarly, we can downsample from the posterior to get a sense of our inferred serology across age groups. 

```{r, fig.width=8, fig.height=6, fig.align='center'}
serocurve <- COVIDCurve::draw_posterior_sero_curves(IFRmodel_inf = modout,
                                                    whichrung = paste0("rung", 1),
                                                    by_chain = F,
                                                    dwnsmpl = 1e3)
#......................
# tidy up posterior draw 
#......................
serocurvedat <- serocurve %>%
  dplyr::select(c("sim", "ObsDay", dplyr::starts_with("RG_pd_"),
                  dplyr::starts_with("crude_pd_"))) %>%
  tidyr::pivot_longer(., cols = -c("sim", "ObsDay"),
                      names_to = "seroprev_strata_lvl", values_to = "seroprev") %>%
  dplyr::mutate(seroprevlvl = ifelse(stringr::str_detect(seroprev_strata_lvl, "RG_"), "RG Corr.", "Crude"),
                param = stringr::str_extract(seroprev_strata_lvl, "ma[0-9]+"))

SeroPrevObs <- tibble::tibble(obsdaymin = sapply(sero_days, median) - 5,
                              obsdaymax = sapply(sero_days, median) + 5)

#......................
# observed
#......................
# quickly add in binomial confidence intervals
inputdata$obs_serology <- inputdata$obs_serology %>% 
  dplyr::mutate(SeroLCI = purrr::map2_dbl(.x = SeroPos, .y =  SeroN,
                                          .f = function(x,y){ binom.test(x,y)$conf.int[1] }),
                SeroUCI = purrr::map2_dbl(.x = SeroPos, .y =  SeroN,
                                          .f = function(x,y){ binom.test(x,y)$conf.int[2] })) %>%
  dplyr::mutate(midpoint = (SeroStartSurvey + SeroEndSurvey)/2)

# curve data
serocurvedat %>%
  ggplot() +
  geom_line(aes(x = ObsDay, y = seroprev, color = seroprevlvl), alpha = 0.5) +
  geom_rect(data = SeroPrevObs, aes(xmin = obsdaymin, xmax = obsdaymax, ymin = -Inf, ymax = Inf),
            fill = "#d9d9d9", alpha = 0.4) +
  geom_pointrange(data = inputdata$obs_serology,
                  aes(x = midpoint, y = SeroPrev, ymin = SeroLCI, ymax = SeroUCI),
                  color = "#000000", alpha = 0.8) +
  facet_wrap(.~param) +
  scale_color_manual("Seroprev. \n Adjustment", values = c("#FFD301", "#246BCF"),
                     labels = c("Inferred 'Truth'", "Inferred 'Observed' - \n Rogan-Gladen Corrected")) +
  labs(caption = "Grey box is observed seroprevalence across study period") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.90, hjust= 1, face = "bold"))


```

#### Posterior Death Check
Finally, we can see how closely our modelled deaths align with the oberserved deaths as a posterior check. 

```{r, fig.width=8, fig.height=6, fig.align='center'}
postdeaths <- COVIDCurve::posterior_check_infxns_to_death(IFRmodel_inf = modout,
                                                          dwnsmpl = 1e3,
                                                          by_chain = FALSE)
postdeaths.plotObj <- postdeaths %>%
  dplyr::select(c("time", dplyr::starts_with("deaths"))) %>%
  tidyr::gather(., key = "Strata", value = "deaths", 2:ncol(.)) %>%
  dplyr::mutate(Strata = gsub("deaths_", "", Strata)) %>% 
  ggplot() +
  geom_line(aes(x= time, y = deaths, group = Strata, color = Strata)) +
  scale_color_viridis_d()

postdeaths.plotObj +
  geom_line(data = simdat$StrataAgg_TimeSeries_Death, aes(x=ObsDay, y = Deaths, group = Strata), color = "#bdbdbd", size = 0.5, alpha = 0.75) +
  facet_wrap(.~Strata) +
  theme_bw() +
  ggtitle("Posterior Predictive Check", subtitle = "Grey Lines are Simulated Data, Viridis Lines are Draws from Posterior")



```
